---
title: "AIDS Drug"
author: "Demian Romero"
date: "11/18/2019"
output:
  html_document:
    number_section: true
    toc: true
---
```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
avast <-read.csv(file="avastivironal.csv",header=TRUE)
avastIV = filter(avast, Ivdrug=='user')
avastGender = filter(avast, sex=='F')
avastLBGT = filter(avast, LBGT=='yes')
head(avast)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative for aids (abacavirinol) is under test in clinical trials. We analyzed the effectiveness of this drug across population subgroups including gender, drug use, and sexual orientation.

The first question this: does abacavirinol have any affect on whether one contracts AIDS or not?

$H_0:$ Abacavirinol has no effect on whether one contracts AIDS or not.  
$H_1:$ Abacavirinol has an effect on whether one contracts AIDS or not.

The second question is:does abacavirinol have a different affect based on gender?

$H_0:$ Gender has no effect on whether abacavirinol offers any protection agains contracting AIDS.

The third question is:does abacavirinol have a different affect based on drug use?

$H_0:$ Drug use has no effect on whether abacavirinol offers any protection agains contracting AIDS.

The final question is:does abacavirinol have a different affect based on sexual orientation?

$H_0:$ sexual orientation has no effect on whether abacavirinol offers any protection against contracting AIDS.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
For each hypothesis test the graphical summary will be a bar plot, the numerical summary will be a cross tabulation, and a chi-squared test will be used for the inferential results.

#Results General Population

is the vaccine effective in the general population?

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->


## Descriptive Results
We use a bar chart to show the effectiveness of the vaccine on the general population.


### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected+treatment, data=avast, type="percent")
```

The barchart shows that the people that received the vaccine are less likely to contract AIDS if they receive the vaccine.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
genTable = xtabs(~infected+treatment, data=avast)
print(genTable)
```
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisqtestGC(genTable)
fisher.test(genTable)
```

This table shows that people whom receive the vaccine are slightly less likely to contract AIDS if they receive the vaccine.

#Results Gender Effects

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
Does gender have an effect on vaccine effectiveness?

## Descriptive Results for Gender

We use a bar chart to show the effectiveness of the vaccine on the general population, allowing for Gender.

### Graphical Descriptive Results for Gender

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected+treatment, data=avastGender, type="percent", xlab='vaccine', main="Gender")
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
xtabs(~infected+treatment, data=avastGender)
```

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~infected+treatment, data=avastGender)
```

From the bar chart, chi-squared test, and numberic table the results appear to demonstrate that the vaccine use results in no measurable difference in infection rates after receiving the vaccine among males versus females.

#Results IV Drug Use

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
Does IV Drug use have an effect on vaccine effectiveness?

## Descriptive Results for IV Drug Use
We use a bar chart to show the effectiveness of the vaccine among IV Drug Users.

### Graphical Descriptive Results for IV Drug Use

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected+treatment, data=avastIV, type="percent", xlab='Vaccine', main='IV Drug')
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
genTable = xtabs(~infected+treatment, data=avastIV)
print(genTable)
```
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~infected+treatment, data=avastIV)
fisher.test(genTable)
```


#Results Sexual Orientation

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

Does sexual orientation have an effect on vaccine effectiveness? 

## Descriptive Results for Sexual Orientation
We use a bar chart to show the effectiveness of the vaccine on the general population allowing for Sexual Orientation.

### Graphical Descriptive Results for Sexual Orientation

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->
```{r}
barchartGC(~infected+treatment, data=avastLBGT, type="percent", xlab='Sexual Orientation')
```

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
genTable2 = xtabs(~infected+treatment, data=avastLBGT)
print(genTable2)
```
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisqtestGC(~infected+treatment, data=avastLBGT)
fisher.test(genTable2)
```


# Final Discussion and Conclusions

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

OVERALL
Null hypothesis is that the vaccine has no effect on the general infected population.  

The p-value was  0,so we reject the null hypothesis which is smaller than .05.
Because the response and explanatory values are categorical, the bar plot and chi-square test for independence.  


GENDER
The results demonstrate that the vaccine is reasonably effective on the general population. 
Null hypothesis is that the vaccine has no effect on the general infected population.

The p-value was 0.3048 , so we fail to reject the null hypothesis which is smaller than .05. Because the response and explanatory values are categorical, the bar plot and chi-square test for independence.


IV DRUG
The results demonstrate that the vaccine is slightly less effective with IV drug users.    
The null hypothesis was that IV Drug use had no effect on the effectiveness of the vaccine.

The p-value was  0.155 , so we fail to reject the null hypothesis which is greater than .05. Because the response and explanatory values are categorical, the bar plot and chi-squared test | test for independence.


LBGT
The results demonstrate that the vaccine is not as effective on the LBGT population.   
The null hypothesis was that sexual orientation does not effect the effectiveness of the vaccine.

The p-value was 0.0104, so we reject the null hypothesis which is smaller than .05. Because the response and explanatory values are categorical, the bar plot and chi-squared test for independence.



The vaccine looks like it is somewhat effective on the general population. It seems to be unaffected by gender or Sexual orientation. The testing indicates we reject the null hypotheses and accept that the vaccine appears to have some positive effect, which is slightly less for some subgroups including IV drug users. 